---
title: "0200-fiss-site-tidy"
date: "Created: 2024-06-20 | Updated: `r format(Sys.Date(), '%Y-%m-%d')`"
output: 
  html_document:
    code_folding: "hide"
params:
  repo_owner: "NewGraphEnvironment"
  repo_name: "fish_passage_template_reporting"
  gis_name: "sern_peace_fwcp_2023"
  job_name: "2024-073-sern-peace-fish-passage"
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=TRUE, include = TRUE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, dpi=60, out.width = "100%", eval = FALSE)
options(scipen=999)
options(knitr.kable.NA = '--') #'--'
options(knitr.kable.NAN = '--')
```


# Purpose of this script

This scripts is used to prepare fiss site data (phase 2) for submission to the province. 

This script is currently being used to clean the 2024 Peace `form_fiss_site_2024.gpkg` file.


# Set up the parameters

To set up the parameter, update the `gis_name` and the `repo_name` in the YAML at the beginning of this script. 
Then update the year in the code below.

```{r params}

path_form_fiss_site <- fs::path('~/Projects/gis/sern_peace_fwcp_2023/data_field/2024/form_fiss_site_2024.gpkg')
```



# Clean the data in preparation for QA in QGIS 

## Fetch and backup the data

The first step is to remove by-hand the first row in `form_fiss_site_2024.gpkg` because it is just a place holder. It has `stream_name` set to `Robert Hatch`

Then we use `fpr_sp_gpkg_backup` to import and backup the data, as well as update the UTMS. We will also write the data to a csv and Rdata file and commit for version control using git.

```{r import}

form_fiss_site_raw <- fpr::fpr_sp_gpkg_backup(
  path_gpkg = path_form_fiss_site,
  dir_backup = "data/backup/",
  update_utm = TRUE,
  update_site_id = FALSE,
  write_back_to_path = FALSE,
  return_object = TRUE,
  col_easting = "utm_easting",
  col_northing = "utm_northing"
) 
```


## Identify duplicate sites

Identify duplicate sites (that are not NAs) as we don't want to input two of the same. 
If there are duplicates go back to QGIS and clean up in `form_fiss_site_2024`. Then read import using scripts above. 

Peace 2024 - No duplicate sites found

```{r duplicates}
dups <- form_fiss_site_raw  |> 
  filter(!is.na(local_name)) |> 
  group_by(local_name) |> 
  filter(n()>1)
```


## Clean up the form 

```{r clean-form}
# clean up the form
form_fiss_site_cleaned <- form_fiss_site_raw |> 
  # split the local_name into the site_id and the location
  tidyr::separate(local_name, into = c('site_id', 'location'), remove = F, extra = "merge") |> 
  # split out the date and the time - change type of column first
  dplyr::mutate(date_time_start = lubridate::ymd_hms(date_time_start),
                date_time_start = lubridate::floor_date(date_time_start, unit = "second"),  # Remove microseconds
                survey_date = lubridate::date(date_time_start),
                time = hms::as_hms(date_time_start), 
                # Fix some vocabulary. Change "trib" to long version "Tributary" etc.
                gazetted_names = str_replace_all(gazetted_names, 'Trib ', 'Tributary '),
                crew_members = toupper(crew_members),
                # fill in text columns from spreadsheet that will likely never change
                waterbody_type = 'stream',
                method_for_utm = 'GPS general',
                method_for_channel_width = 'metre tape',
                method_for_wetted_width = 'metre tape',
                method_for_residual_pool_depth = 'metre stick',
                method_for_bankfull_depth = 'metre stick',
                method_for_gradient = 'clinometer',
                method_for_temperature = 'recording meter',
                method_for_conductivity = 'recording meter',
                method_for_p_h = 'pH meter (general)') |> 
  # arrange by surveyor and date/time
  dplyr::arrange(mergin_user, date_time_start) |> 
  dplyr::mutate(comments = paste0(comments, time)) |> 
  # ditch the time since we don't need anymore. Time was dropped on gpkg creation due to type conflict
  select(-time) |>
  # rearrange the columns for easier QA in QGIS.
  dplyr::select(
    date_time_start,
    local_name,
    gazetted_names,
    crew_members,
    comments,
    everything()) |> 
  arrange(date_time_start)

```


## Burn a cleaned copy to the QGIS project for QA

```{r burn-form-fiss-site}
# Burn cleaned copy to the QGIS project
form_fiss_site_cleaned |> 
  sf::st_write(path_form_fiss_site, append=FALSE, delete_dsn = T)
```

Now open QGIS and re-add the `form_fiss_site_2024` layer. Check to see the form is updated with the changes made here. Copy/paste all
`styles` from the old layer to the new layer you just added.  Lastly, you can delete the old layer and start QAing `form_fiss_site_2024.` 
