---
title: "0300-monitoring-backup-wrangle"
date: "Created: 2025-03-10 | Updated: `r format(Sys.Date(), '%Y-%m-%d')`"
output: 
  html_document:
    code_folding: "hide"
params:
  repo_owner: "NewGraphEnvironment"
  repo_name: "fish_passage_skeena_2024_reporting"
  gis_project_name: "sern_skeena_2023"
  job_name: "2024-072-sern-skeena-fish-passage"
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=TRUE, include = TRUE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, dpi=60, out.width = "100%", eval = FALSE)
options(scipen=999)
options(knitr.kable.NA = '--') #'--'
options(knitr.kable.NAN = '--')
```


# Purpose of this script

This scripts is used to backup the `form_monitoring_2024.gpkg` file before QA in QGIS and to clean the form after QA in QGIS.


## Fetch and backup the data

Update the `gis_project_name` and the `repo_name` in the YAML at the beginning of this script. 

Name the project directory we are pulling from and import the data. We use `fpr_sp_gpkg_backup` to import and backup the data, as well as update the UTMS. We will also write the data to a csv and Rdata file and commit for version control using git. Commit these inital backup files with a message such as "initial commit - raw 2024 data".

`update_site_id = TRUE` now checks for duplicate sites.

```{r import}
path_form_monitoring <- fs::path_expand(fs::path("~/Projects/gis/", params$gis_project_name, "/data_field/2024/form_monitoring_2024.gpkg"))

form_monitoring_raw <- fpr::fpr_sp_gpkg_backup(
  path_gpkg = path_form_monitoring,
  dir_backup = "data/backup/",
  update_utm = TRUE,
  update_site_id = FALSE, ## This now also checks for duplicates
  write_back_to_path = FALSE,
  write_to_csv = TRUE,
  write_to_rdata = TRUE,
  return_object = TRUE,
  col_easting = "utm_easting",
  col_northing = "utm_northing")
  
```


## QA in QGIS

The following should be done in QGIS:

- remove the site used to make the form (with `my_crossing_reference` set to `123456`)
- back up the original assessment comments so we can redo this amalgamation of text if we need to. Use the field calculator, select generate new field, name it `assessment_comment_og`, and select `assessment_comment` in the expression box. 
- reorder the columns for easy QA in Q


## Backup the data after QA

Backup the data after QA in QGIS and commit with a detailed message.

```{r backup}
path_form_monitoring <- fs::path_expand(fs::path("~/Projects/gis/", params$gis_project_name, "/data_field/2024/form_monitoring_2024.gpkg"))

form_monitoring <- fpr::fpr_sp_gpkg_backup(
  path_gpkg = path_form_monitoring,
  dir_backup = "data/backup/",
  update_utm = TRUE,
  update_site_id = TRUE, ## This now also checks for duplicates
  write_back_to_path = FALSE,
  write_to_csv = TRUE,
  write_to_rdata = TRUE,
  return_object = TRUE,
  col_easting = "utm_easting",
  col_northing = "utm_northing")
  
```



## If you need to load an old version of a form

This is code to load an an old old version of a form and reburn to a geopackage 

```{r load-from-backup}

# Load the backup .Rdata file for the old version of the form
load("/Users/lucyschick/Projects/current/temp_fiss_backup/form_fiss_site_2024.RData")

form_fiss_old <- form_fiss_site_2024
rm(form_fiss_site_2024)

# Burn back to a geopackage
form_fiss_old |> 
  sf::st_write(path_form_fiss_site,
               append = FALSE,
               delete_dsn = TRUE)
```

